### 2020-04-28 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font

%\usepackage{mathpazo} % Palatino font
%\renewcommand*\familydefault{\rmdefault} % Use the roman version of the font

\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{ten-simple-rules-to-increase-computational-skills-among-biologists-with-code-clubs}{%
\section{Ten Simple Rules to increase computational skills among
biologists with Code
Clubs}\label{ten-simple-rules-to-increase-computational-skills-among-biologists-with-code-clubs}}

\vspace{5mm}

Ada K. Hagan\({^{1,2}}\), Nicholas A. Lesniak\({^1}\), Marcy J.
Balunas\({^3}\), Lucas Bishop\({^1}\), William L. Close\({^1}\), Matthew
D. Doherty\({^1}\), Amanda G. Elmore\({^1}\), Kaitlin J.
Flynn\({^{1,4}}\), Geoffrey D. Hannigan\({^{1,5}}\), Charlie C.
Koumpouras\({^{1,6}}\), Matthew L. Jenior\({^{1,7}}\), Ariangela J.
Kozik\({^8}\), Kathryn McBride\({^1}\), Samara B. Rifkin\({^8}\), Joshua
M.A.~Stough\({^1}\), Kelly L. Sovacool\({^9}\), Marc A.
Sze\({^{1,10}}\), Sarah Tomkovich\({^1}\), Begum D.
Topcuoglu\({^{1,5}}\), and Patrick D. Schloss\({^1}\)\({^\dagger}\)

\vspace{5mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{pschloss@umich.edu}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Department of Microbiology and Immunology, University of Michigan, Ann
  Arbor, MI
\item
  Current affiliation: Alliance SciComm \& Consulting, Linden, MI
\item
  Division of Medicinal Chemistry, Department of Pharmaceutical
  Sciences, University of Connecticut, Storrs, CT
\item
  Current affiliation: Benaroya Research Institute, Seattle, WA
\item
  Current affiliation: Exploratory Science Center, Merck \& Co., Inc.,
  Cambridge, Massachusetts, USA
\item
  Current affiliation: Roche Diagnostics, Clinical Operations Services
  \& eSystems, Indianapolis, IN
\item
  Current affiliation: Department of Biomedical Engineering, University
  of Virginia, Charlottesville, VA
\item
  Department of Internal Medicine, University of Michigan
\item
  Department of Computational Medicine \& Bioinformatics, University of
  Michigan
\item
  Current affiliation: Genomics and Pharmacogenomics, Merck \& Co.,
  Inc., Cambridge, MA
\end{enumerate}

\vspace{5mm}

\hypertarget{ten-simple-rules}{%
\subsubsection{Ten Simple Rules}\label{ten-simple-rules}}

\newpage
\linenumbers

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

For most biologists, the ability to generate data has outpaced the
ability to analyze those data. High throughput data comes to us from DNA
and RNA sequencing, flow cytometry, metabolomics, molecular screens, and
more. Although some accept the approach of compartmentalizing data
generation and data analysis, we have found scientists feel empowered
when they can both ask and answer their own biological questions
{[}1{]}. Yet, the standard undergraduate and graduate training in the
biological sciences is typically insufficient for developing these
advanced data analysis skills. In our experience performing microbiome
research, it is more common to find exceptional bench scientists who are
inexperienced at analyzing large data sets than to find the reverse. Of
course this raises a challenge: How do we train bench scientists to
effectively answer biological questions with these larger datasets?

The ever-growing ability to generate data and feeling of unease in
analyzing the data are analogous to the struggles scientists also face
with engaging the voluminous scientific literature. A common strategy
for keeping up with the literature is participating in journal clubs,
which involve group discussion of a pre-selected paper. These range from
informal discussions to PowerPoint presentations and course credit. In
addition to staying current on the literature, journal clubs help
strengthen skills in critical thinking, communication, and integrating
the literature {[}2{]}. Our research group has leveraged the similarity
between the overwhelming nature of both scientific literature and data
analysis to address the challenge of teaching reproducible data analysis
practices to bench scientists. Over the past four years we have
experimented with a Code Club model to improve data analysis skills in a
community environment.

Our Code Club sessions are an hour long and alternate with Journal Club
as the second part of weekly two hour lab meetings. Initially, the Code
Club was used to review code from trainee projects. Instead of a
presentation, the presenter would project their code onto a screen and
the participants would go through the code, stating the logic behind
each line. This approach emphasized the importance of code readability
and gave beginners the opportunity to see the real-life, messy code of
more experienced peers. Unfortunately, the format only allowed us to
review a fraction of a project's code, making it difficult to integrate
the programmer's logic across their full project. A major issue with
this model was that sometimes beginners could not contribute to
improving the code, and even when they could, more experienced group
members would eventually dominate the discussion. This led to a lack of
participation by beginners, who would sometimes mentally check out and
resulted in an adversarial environment between the presenter and more
experienced members. As a result, presenters were reluctant to offer to
present again.

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}3{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using either R syntax from the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. Our Code Clubs
typically have 7 to 10 participants, but the inherent
``think-pair-share'' strategy should allow it to be scaled to larger
groups {[}4{]}.

We continue to experiment with approaches for running Code Club. During
the period of the Covid-19 pandemic in 2020 when many research labs were
shuttered, PDS started hosting an open virtual weekly Code Club.
Participants could engage the material posted to
\url{https://www.riffomonas.org/code_club/} via a Zoom meeting or
asynchronously using the recorded sessions available on YouTube.
Regardless of our approach to leading a Code Club, we have learned that
it is critical for the presenter to clearly articulate their goals and
facilitate participant engagement. Although some Code Club sessions may
be more experimental than others, on the whole they are a critical tool
to train bench scientists in reproducible data analysis practices. We
have provided some examples of successful topics in Table 1. We have
summarized the results of our own experimentation as Code Club
presenters and participants into Ten Rules. The first 3 rules apply to
all participants, Rules 4 through 8 are targeted to presenters, and the
last two for participants.

\hypertarget{rule-1-reciprocate-respect}{%
\subsection{Rule 1: Reciprocate
respect}\label{rule-1-reciprocate-respect}}

It is critical that the presenter and participants respect each other
and that a designated individual (e.g.~the lab director) enforces a code
of conduct. Each member of the Code Club must have the humility to
acknowledge that they have more to learn about any given topic. We have
found that many problems are avoided when the presenter takes charge of
the session with a clear lesson plan, thoughtfully creates groups, and
gives encouragement. Similarly, participants foster a positive
environment by remembering that the task is not a competition, instead
focusing on the presenter's goals, allowing their partner to contribute,
asking clarifying questions when appropriate, and avoiding distractions
(e.g.~email, social media). Learning to program is challenging. Too
often it is attempted in a environment with nonconstructive criticism.
All parties in a Code Club are responsible for preventing this by
demonstrating respect for themselves and their colleagues.

\hypertarget{rule-2-let-the-material-change-you}{%
\subsection{Rule 2: Let the material change
you}\label{rule-2-let-the-material-change-you}}

Part of the humility required to participate in a Code Club is
acknowledging that your training is incomplete and that it is possible
for everyone to learn something new. For participants, assume that the
presenter has a plan and follow their presented approach. After the Code
Club, try to incorporate that material into new code or refactor old
code. By practicing the material in a different context, you will learn
the material better. For presenters, incorporate suggested changes into
your code. Either party may identify concepts that they are unsure of,
presenting opportunities for further conversation and learning.

\hypertarget{rule-3-experiment}{%
\subsection{Rule 3: Experiment!}\label{rule-3-experiment}}

The selection of content and structure for each Code Club works best
when it is democratic and distributed. If someone thinks a technique is
worth learning and wants to teach it, they have that power. If they want
to experiment with a different format, they are free to try it out.
Members of the Code Club need to feel like they have the power to shape
the direction of the group. If members are following these Rules, they
will naturally reflect on the skills and interests of the other members
in the group. For example, there is always turnover in a research group,
making it important to revisit basic concepts to teach to new people and
provide a refresher to others. Keeping an online log of previous topics
or topics of interest for future sessions is a useful organizational
tool. The group should also feel free to experiment with the format and
incorporate group feedback by ending with a debrief discussing the pros
and cons of new formats.

\hypertarget{rule-4-set-specific-goals}{%
\subsection{Rule 4: Set specific
goals}\label{rule-4-set-specific-goals}}

In our early Code Clubs we noticed that if the presenter did not clearly
state their goals for the session, it often led to frustration for both
the presenter and participants. If the presenter shared their own code,
did they want participants to focus on their coding style or did they
want help incorporating a new package into their workflow? Participants
will always notice or ask about code concepts that are not the focus of
the exercise; a presenter with a specific goal can bring tangential
conversations back to the planned task. Where possible, presenters
should create a simplified scenario (i.e.~minimal, reproducible example
{[}5{]}), which can be helpful in focusing the participants. The
presenter should verify ahead of time that the simplified example works
and behaves the way they expect. Beyond the content, clear goals for
participant activities will help both parties stay on task and avoid
frustration. For more advanced learners, the presenter can create
stretch goals or give an activity with multiple stopping points where
participants would feel successful (e.g.~commenting code, creating
function, implementing function, refactoring function). Accomplishing
specific goals is more likely to result in a positive outcome for
presenters and participants.

\hypertarget{rule-5-keep-it-simple}{%
\subsection{Rule 5: Keep it simple}\label{rule-5-keep-it-simple}}

Our Code Club needs to fit within an hour time slot. When considering
their Code Club activity, the presenter should plan for an introduction
and brief instruction, time for participants to engage the material, and
time for everyone to report back within that hour. A typical schedule
for Code Club is 10 minutes of introduction and instruction, 30 minutes
of paired programming, 5 minutes to get groups to wrap up, and 10
minutes to report back to the group. We once had a presenter try to
teach basic, but unfamiliar, Julia syntax. Unfortunately, the time was
up before the participants had installed the interpreter. Some tips to
help keep it simple are to limit the presented code chunks to less than
50 lines or, conversely, consider the number of lines that might be
required to accomplish a solution. Remember that learners may need up to
three times as long to complete a task that is simple for the presenter,
so Code Club is best kept simple.

\hypertarget{rule-6-give-participants-time-to-prepare}{%
\subsection{Rule 6: Give participants time to
prepare}\label{rule-6-give-participants-time-to-prepare}}

Similar to a Journal Club, the presenter should give participants a few
days (ideally a week) to prepare for the Code Club. Considering the
compressed schedule described in Rule 5, asking participants to download
materials beforehand is helpful to ensure a quick start. The presenter
should provide the participants with instructions on how to install
dependencies, download data, and get the initial code. This might also
uncover weak points in the presenter's plan and enable them to ensure
that the materials work as intended before the Code Club. We have found
that using GitHub repositories for each Code Club can help make
information, scripts, and data easily available to participants. Perhaps
a first Code Club could introduce using git and GitHub to engage in
collaborative coding. A lower barrier entry point for the presenter is
posting their code, data, and information in a lab meeting-dedicated
Slack channel or via email. Whatever method is used, the presenter
should be sure to communicate the topic and necessary materials with the
participants ahead of time.

\hypertarget{rule-7-dont-give-participants-busy-work}{%
\subsection{Rule 7: Don't give participants busy
work}\label{rule-7-dont-give-participants-busy-work}}

Participants want to learn topics that will either be useful to them or
help their colleague (i.e.~the presenter). Presenters should do their
best to satisfy those motivations, whether it is through the relevancy
of the concept or the data. It does not make sense to present a Code
Club on downloading stock market data if it is not useful or interesting
to the group. Similarly, participants should not be tasked with
improving the presenter's code if the presenter has no intention of
incorporating the suggestions. A list of packages or tasks that group
members are interested in could help a presenter struggling to find a
topic choose one that will make a rewarding Code Club.

\hypertarget{rule-8-include-all-levels-of-participants}{%
\subsection{Rule 8: Include all levels of
participants}\label{rule-8-include-all-levels-of-participants}}

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. In these circumstances it can be useful to cover commonly
used tools but that might result in disinterest of more experienced
participants. We have identified several strategies to overcome the
challenges presented by participants at varying skill levels. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on their goals. Partnerships between those with similar skill
levels requires the presenter to design appropriate activities for each
skill level. We have found that commenting code is a good skill for
beginners since it forces them to dissect and understand a code chunk
line-by-line. It also reinforces the value of commenting as they develop
their skills and independence. An advantage of forming partnership
between people with disparate skill levels is that it is more likely for
groups to provide the presenter with a diverse range of methods that
achieve the same result. This approach to pairing also helps to graft
new members that have emerging programming skills. Regardless of how
partners are selected, consider asking the pairs to identify a navigator
and a driver {[}6{]}. The driver types at the computer while the
navigator tells them what to type, thus ensuring participation of both
partners. Midway through the activity, the presenter can have the
partners switch roles. Intentionally forming pairs can also engineer
group interactions by avoiding potentially disruptive partnerships or
pairing reliable role models with new group members.

\hypertarget{rule-9-prepare-in-advance-to-maximize-participation}{%
\subsection{Rule 9: Prepare in advance to maximize
participation}\label{rule-9-prepare-in-advance-to-maximize-participation}}

It is not possible to fully participate in a Journal Club discussion
about a paper that the participant has not read. In that context, coming
to Code Club without having installed a dependency is similar to asking
a simple question about the Journal Club paper. Both instances show a
lack of preparation. Just as a presenter must follow Rule 6 to provide
materials ahead of time, participants must review the code in advance,
download the data sets, install the necessary packages, and perhaps read
up on the topic. If the Code Club is based on a paper or chapters in a
data science book (e.g., {[}7--9{]}), the participants should read them
before the session and consider how they might incorporate the concepts
into their own work.

\hypertarget{rule-10-participate}{%
\subsection{Rule 10: Participate}\label{rule-10-participate}}

An essential ingredient of any Code Club is active participation from
all parties. Having an open laptop on the table and permission to use it
can feel like an invitation to get distracted by other work, emails, and
browsing the internet; fight that urge and focus on the presenter's
goals. Be respectful and allow your partner to contribute. Speak up for
yourself and force your partner to let you contribute. If the material
seems too advanced for you, it can be frustrating, and tempting to
mentally check out. Fortunately, programming languages like R and Python
are generally expressive, which should allow you to engage with the
logic, even if the syntax is too advanced. Oftentimes, understanding the
logic of \emph{when} to use one modeling approach over another is more
important than knowing \emph{how} to use it. If you understand the
``why'', the ``how'' will quickly follow. More experienced participants
should aim to communicate feedback and coding suggestions at a level
that all participants can understand and engage in. Regardless of skill
levels, your partner and the presenter put themselves in a vulnerable
position by revealing what they do or do not know. Encourage them, and
show your gratitude for helping you learn something new, by fully
participating in each Code Club.

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}

The most important rules are the first and last. Members of the Code
Club need to feel comfortable with other group members and sufficiently
empowered to try something new or ask for help. Aside from expanding our
programming skills, we have noticed two other benefits that help create
a positive work culture. First, we have intentionally interviewed
postdoc candidates on the days we hold Code Club. We make it clear that
they are not being assessed on their coding skills, but instead use it
as an opportunity to see how the candidate interacts with other members
of the research group. At the same time, the candidate can learn about
the culture of the research group through active participation. Second,
members of other research groups have integrated themselves into our
Code Club to minimize the isolation they feel in growing their skills
within smaller research groups. This speaks to both the broader need for
Code Club and the likelihood of success when expanded to include a
larger group of individuals with broader research interests. Ultimately,
Code Club has improved the overall data analysis skills, community, and
research success of our lab by empowering researchers to seek help from
their colleagues.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

All co-authors have been participants in the Code Clubs run as part of
PDS's lab meetings. They have had a critical role in shaping the
evolution of the sessions. AKH led a lab discussion of Code Clubs to
identify the key concepts presented, which were supplemented by
contributions from NAL. The order of other co-authors is listed
alphabetically.

\newpage

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Carey2018}{}%
1. Carey MA, Papin JA. Ten simple rules for biologists learning to
program. Markel S, editor. PLOS Computational Biology. 2018;14:
e1005871.
doi:\href{https://doi.org/10.1371/journal.pcbi.1005871}{10.1371/journal.pcbi.1005871}

\leavevmode\hypertarget{ref-Lonsdale2016}{}%
2. Lonsdale A, Penington JS, Rice T, Walker M, Dashnow H. Ten simple
rules for a bioinformatics journal club. Lewitter F, editor. PLOS
Computational Biology. 2016;12: e1004526.
doi:\href{https://doi.org/10.1371/journal.pcbi.1004526}{10.1371/journal.pcbi.1004526}

\leavevmode\hypertarget{ref-Wilson2014}{}%
3. Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et al.
Best practices for scientific computing. Eisen JA, editor. PLoS Biology.
2014;12: e1001745.
doi:\href{https://doi.org/10.1371/journal.pbio.1001745}{10.1371/journal.pbio.1001745}

\leavevmode\hypertarget{ref-Kothiyal2014}{}%
4. Kothiyal A, Murthy S, Iyer S. Think-pair-share in a large CS1 class.
Proceedings of the 2014 conference on innovation \& technology in
computer science education - ITiCSE 14. ACM Press; 2014.
doi:\href{https://doi.org/10.1145/2591708.2591739}{10.1145/2591708.2591739}

\leavevmode\hypertarget{ref-StackOverflow_MREE}{}%
5. Stack Overflow. How to create a minimal, reproducible example. 2020.
Available:
\url{https://stackoverflow.com/help/minimal-reproducible-example}

\leavevmode\hypertarget{ref-Bockeler_2020}{}%
6. Böckeler B, Siessegger N. On pair programming. 2020. Available:
\url{https://martinfowler.com/articles/on-pair-programming.html}

\leavevmode\hypertarget{ref-Wickham_2016}{}%
7. Wickham H, Grolemund G. O'Reilly Media; 2016. Available:
\url{https://r4ds.had.co.nz}

\leavevmode\hypertarget{ref-Schloss_2018}{}%
8. Schloss PD. The riffomonas reproducible research tutorial series.
Journal of Open Source Education. 2018;1: 13.
doi:\href{https://doi.org/10.21105/jose.00013}{10.21105/jose.00013}

\leavevmode\hypertarget{ref-Noble2009}{}%
9. Noble WS. A quick guide to organizing computational biology projects.
Lewitter F, editor. PLoS Computational Biology. 2009;5: e1000424.
doi:\href{https://doi.org/10.1371/journal.pcbi.1000424}{10.1371/journal.pcbi.1000424}

\newpage

\textbf{Table 1. Examples of successful Code Club topics.}

\begin{longtable}[]{@{}ll@{}}
\toprule
\begin{minipage}[b]{0.25\columnwidth}\raggedright
\textbf{Title}\strut
\end{minipage} & \begin{minipage}[b]{0.69\columnwidth}\raggedright
\textbf{Description}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.25\columnwidth}\raggedright
base vs.~ggplot2\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given input data and a figure, recreate the figure using R's base
graphics or ggplot2 syntax\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Snakemake\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a bash script that contains an analysis pipeline, convert it to a
Snakemake workflow (can also be done with GNU Make)\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
DRYing code\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given script with repeated code, create functions to remove
repetition\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
mothur and Vegan\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a pairwise community dissimilarity matrix, compare communities
using the \texttt{adonis} function in the Vegan R package\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
tidy data\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Given a wide-formatted data table, convert it to a long, tidy-formatted
data table using tools from R's tidyverse\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
GitFlow\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Participants file and claim an issue to add their name to a README file
in a GitHub-hosted repository and file a pull request to complete the
issue\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
R with Google docs\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Scrape a Google docs workbook and clean the data to identify previous
Code Club presenters\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.25\columnwidth}\raggedright
Develop an R Package\strut
\end{minipage} & \begin{minipage}[t]{0.69\columnwidth}\raggedright
Convert a lab member's collection of scripts into an R package over a
series of sessions\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\end{document}